Coherent optical association of a single molecule

### 2020-11-04 paper.tex
Font

The essence of an optical Raman transfer can be illustrated using a three-level system (Fig.~\ref{f-theory}a), where the initial atomic state, and the target weakly-bound molecular state are coupled to an intermediate state by two laser  with Rabi frequencies, $\Omega_a$ and $\Omega_m$, with one-photon detuning $ \Delta $, and two-photon detuning, $ \delta$.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $c^3\Sigma^+$)
The transfer Raman Rabi Rate, $\Omega_R=\Omega_a\Omega_m / 2\Delta$, is accompanied by a photon scattering rate $\Gamma_s$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
~\cite{Wineland2003}.
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and therefore the scattering is predominantly from the target molecular state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single photon detuning, $ \Delta $, so the target molecular state can scatter off both beams roughly equally. Thus, the scattering is given by $ \Gamma_s=\Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth \footnote{We choose the two beams to have equal power, which gives the highest Raman Rabi rate at a fixed total power. Thus, this results in a simple factor of 2 coming from scattering off 2 beams.}.
The ratio between the Raman Rabi frequency and the scattering rate is therefore $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
To ensure a coherent process, a detuning as large as possible, while maintaining a realistic Raman Rabi frequency, is preferred. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

To find the optimal intermediate state, we perform a calculation of the Raman Rabi frequency
and scattering rate at different detunings from the atomic threshold
taking into account of all states of
the $c^3\Sigma^+(\Omega = 1)$ excited molecular state potential~\cite{Grochola2011}
and the continuum~\cite{Liu2017}.
The excited atomic continuum is particularly important for the target molecular state.
The sum of the squares of the wavefunction overlap
between the target weakly-bound molecular state
and all the excited molecular bound states is only about $0.02$,
suggesting that there is significant matrix element
between the target molecular state and the excited atomic continuum (See SM \todo{proper cite}).
This calculation shows that the ratio of the Raman Rabi rate to scattering rate
can be made larger for more deeply bound states compared to weakly-bound states
at a cost of a smaller Raman Rabi frequency
(Fig.~\ref{f-theory}b, full result in SM \todo{proper cite})
As a result, we choose the $v'=0$ of $c^3\Sigma^+(\Omega = 1)$ as an intermediate state
to drive the Raman transition.

@@

In conclusion, we have formed a weakly-bound NaCs molecule in an optical tweezer
via an optical Raman transfer.
A theoretical investigation including all excited states of $ c^3\Sigma^+(\Omega = 1)$,
the excited atomic continuum, and coupled-channel ground state wavefunctions indicated
better transfer efficiency using a deeply bound intermediate state
and the $\ket{\uparrow_{\Na} \downarrow_{\Cs} }$ hyperfine state as the initial and final states.
Using these theoretical insights, we located the weakly-bound state
and coherently associated the atoms into the weakly-bound molecule.
Our transfer efficiency is limited by an unknown scattering source
resulting in measured scattering rates over $10$ times larger than theoretical predictions.
Despite this limitation, the transfer efficiency may be further improved
by increasing the ratio of the up-leg to down-leg Rabi frequency
$\Omega_a/\Omega_m$ by exploring the possibility of
driving to more deeply bound states.
There may also be a better choice of single photon detuning to increase the Raman Rabi frequency,
since our location results in about $30~\mathrm{\%}$ cancellation of the Raman Rabi frequency
due to an offset of unknown origin.

>>>>

The essence of an optical Raman transfer can be illustrated using a three-level system (Fig.~\ref{f-theory}a), where the initial atomic state, and the target weakly-bound molecular state are coupled to an intermediate state by two laser  with Rabi frequencies, $\Omega_a$ and $\Omega_m$, with one-photon detuning $ \Delta $, and two-photon detuning, $ \delta$.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi Rate, $\Omega_R=\Omega_a\Omega_m / 2\Delta$, is accompanied by a photon scattering rate $\Gamma_s$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
~\cite{Wineland2003}.
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and therefore the scattering is predominantly from the target molecular state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single photon detuning, $ \Delta $, so the target molecular state can scatter off both beams roughly equally. Thus, the scattering is given by $ \Gamma_s=\Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth \footnote{We choose the two beams to have equal power, which gives the highest Raman Rabi rate at a fixed total power. Thus, this results in a simple factor of 2 coming from scattering off 2 beams.}.
The ratio between the Raman Rabi frequency and the scattering rate is therefore $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
To ensure a coherent process, a detuning as large as possible, while maintaining a realistic Raman Rabi frequency, is preferred. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

To find the optimal intermediate state, we perform a calculation of the Raman Rabi frequency
and scattering rate at different detunings from the atomic threshold
taking into account of all states of
the $\mathrm{c^3\Sigma^+}(\Omega = 1)$ excited molecular state potential~\cite{Grochola2011}
and the continuum~\cite{Liu2017}.
The excited atomic continuum is particularly important for the target molecular state.
The sum of the squares of the wavefunction overlap
between the target weakly-bound molecular state
and all the excited molecular bound states is only about $0.02$,
suggesting that there is significant matrix element
between the target molecular state and the excited atomic continuum (See SM \todo{proper cite}).
This calculation shows that the ratio of the Raman Rabi rate to scattering rate
can be made larger for more deeply bound states compared to weakly-bound states
at a cost of a smaller Raman Rabi frequency
(Fig.~\ref{f-theory}b, full result in SM \todo{proper cite})
As a result, we choose the $v'=0$ of $\mathrm{c^3\Sigma^+}(\Omega = 1)$ as an intermediate state
to drive the Raman transition.

@@

In conclusion, we have formed a weakly-bound NaCs molecule in an optical tweezer
via an optical Raman transfer.
A theoretical investigation including all excited states of $\mathrm{c^3\Sigma^+}(\Omega = 1)$,
the excited atomic continuum, and coupled-channel ground state wavefunctions indicated
better transfer efficiency using a deeply bound intermediate state
and the $\ket{\uparrow_{\Na} \downarrow_{\Cs} }$ hyperfine state as the initial and final states.
Using these theoretical insights, we located the weakly-bound state
and coherently associated the atoms into the weakly-bound molecule.
Our transfer efficiency is limited by an unknown scattering source
resulting in measured scattering rates over $10$ times larger than theoretical predictions.
Despite this limitation, the transfer efficiency may be further improved
by increasing the ratio of the up-leg to down-leg Rabi frequency
$\Omega_a/\Omega_m$ by exploring the possibility of
driving to more deeply bound states.
There may also be a better choice of single photon detuning to increase the Raman Rabi frequency,
since our location results in about $30~\mathrm{\%}$ cancellation of the Raman Rabi frequency
due to an offset of unknown origin.
